Expert-In-The-Loop Causal Discovery: An Iterative Approach for Efficiently Utilizing Expert Knowledge for Model Modification

### 2024-08-06 paper/main.tex
Updates

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Expert-In-The-Loop Causal Discovery}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

\begin{abstract}
	Causal Discovery has received substantial attention in the DAG
	literature and many automated algorithms have been proposed for
	constructing DAGs from data. However, the application of these
	algorithms in applied fields remain limited due to challenges in using
	them in practice for example they make obvious mistakes, output CPDAGs
	that need to be oriented using expert knowledge before estimation
	methods can be applied. As a result of these challenges, researchers
	typically construct these model by hand. In this paper, inspired by
	specification search in SEMs where the method suggests changes in the
	model that improves the fit the most, we present a method based on
	Conditional Independence testing and a novel measure of partial
	association. The method iteratively suggests addition and removal of
	edges between pairs of variables, while the user can choose the
	direction of the edge. Using empirical analysis we show that this
	manual construction approach can outperform automated algorithms if the
	user gets the direction correct in <x out of y> cases. We also show
	that the user expert can be replaced with an LLM.
\end{abstract}

@@

\begin{enumerate}
	\item \textbf{Make Obvious Mistakes:} Although most algorithms are
		theoretically consistent asymptotically, their finite sample
		properties are not well understood. Moreover, depending on the
		selected hyperparameters, their output can vary significantly.
		These issues combined with no good method to test the
		performance of these algorithms on a given dataset, in
		real-world settings with finite samples, it can be difficult to
		trust the output of these algorithms. Consequently, researchers
		have to rely on posthoc modifications to the model based on
		their domain expertise.
	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple
		DAGs can be faithful to a given observational dataset. Thus,
		using only observational data, algorithms can only recover the
		MECs that can contain a combination of directed and undirected
		edges. However, most methods for downstream tasks, such as
		identification or causal effect estimation, assume knowledge of
		a fully oriented DAG. As a result, researchers have to manually
		orient these edges before using them for analysis.
\end{enumerate}

>>>>

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\title{Expert-In-The-Loop Causal Discovery: An Iterative Approach for Efficiently Utilizing Expert Knowledge for Model Modification}
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,\\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz\equalcontrib,
    Marc Pujol-Gonzalez\equalcontrib
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    % If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    % For example,

@@

\begin{abstract}
	\emph{Causal Discovery} has received significant attention in the
	\emph{Directed Acyclic Graphs} (DAGs) literature, leading to the
	development of numerous automated algorithms for constructing DAGs from
	data. Despite these advancements, the application of these algorithms
	in applied domains remain limited, with researchers often preferring to
	construct DAGs manually based on their expert knowledge. We hypothesize
	that the reluctance to use automated algorithms arises due to several
	practical issues, such as algorithms often make obvious mistakes making
	it difficult to trust, they output the Markov Equivalence Class (MEC)
	instead of a DAG (which is often required for further analysis), and
	difficulties in choosing the appropriate algorithm and hyperparameters
	for a given dataset. Even when using these algorithms, researchers must
	rely on their expert knowledge to make modifications and orient edges
	of the learned model. Given this reliance on manual model modification,
	in this paper, we present a method inspired by \emph{specification
	search} in \emph{Structural Equation Models} (SEMs) to assist
	researchers in iteratively modifying their models to improve their fit
	to data. Specifically, we utilize implied Conditional Independence (CI)
	based model testing combined with a novel measure of partial
	association for mixed data to rank violations of implied CIs of the
	model in data. Using this ranking users can choose to make
	modifications to their models by either removing edges or adding edges
	between pair of variables while using their expert knowledge to decide
	the orientation of the edges. The ranking allows users to prioritize
	the worst offending violations first. Through empirical analysis, when
	applied to learning the model from scratch, we demonstrate that taking
	a greedy approach (i.e., fixing the worst violation first) this
	approach can get comparable performance to automated algorithms if the
	users are able to decide the edge direction correctly with an accuracy
	higher than $ 0.4 $. (TODO: Double check this value). We also show that
	a Large Language Model (LLM) can also be used for the purpose of
	deciding the edge orientation in the absence of an expert.
\end{abstract}

@@

\begin{enumerate}
	\item \textbf{Make Obvious Mistakes:} Although most algorithms are
		theoretically consistent asymptotically, their finite sample
		properties are not well understood. Moreover, depending on the
		selected hyperparameters, their output can vary significantly.
		These issues combined with no good method to test the
		performance of these algorithms on a given dataset, in
		real-world settings with finite samples, it can be difficult to
		trust the output of these algorithms. Consequently, researchers
		have to rely on posthoc modifications to the model based on
		their domain expertise.
	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple
		DAGs can be faithful to a given observational dataset. Thus,
		using only observational data, algorithms can only recover the
		MECs that can contain a combination of directed and undirected
		edges. However, most methods for downstream tasks, such as
		identification or causal effect estimation, assume knowledge of
		a fully oriented DAG. As a result, researchers have to manually
		orient these edges before using them for analysis.
	\item \textbf{Difficulty in Choosing Best Algorithm and Hyperparameters:}
\end{enumerate}
